Learning from Causation

### 2016-03-22 talks/templates/tum-beamer/cbn2012.tex
add talks

%% MAKE TITLE
\title{Basic Beamer Template}
\subtitle{A subtitle should be placed here}
\author{Huang Xiao} %\\ \small{xiaohu@in.tum.de} }%\thanks{Funded by ARAMiS project}}
%\and Author2 \inst{1} }
%\and Author3 \inst{2}}
\institute[Technische Universit\"at M\"unchen]
{ 
	%\inst{1} 
	Chair of IT Security (I20) \\Department of Informatics \\Technische Universit\"at M\"unchen 
	%\and
	%\inst{2} Cloud Service Lab\\Fraunhofer AISEC
}

@@

%% START PRESENTATION
\section{Introduction}

\begin{frame}{Introduction}

\begin{itemize}
\item Define the problem being studied.
\item Explain your interest in the problem.
\item Outline how the problem is to be investigated.
\end{itemize}

\end{frame}
\subsection{XXX1}
\begin{frame}{xxx1}

\begin{itemize}
\item Define the problem being studied.
\item Explain your interest in the problem.
\item Outline how the problem is to be investigated.
\end{itemize}

\end{frame}
\subsection{XXX2}
\begin{frame}{xxx2}

\begin{itemize}
\item Define the problem being studied.
\item Explain your interest in the problem.
\item Outline how the problem is to be investigated.
\end{itemize}

\section{History}

>>>>

%% MAKE TITLE
\title{Learning from Causation}
\subtitle{Fundamental of Casual Inference and its Applications}
\author{Huang Xiao} %\\ \small{xiaohu@in.tum.de} }%\thanks{Funded by ARAMiS project}}
%\and Author2 \inst{1} }
%\and Author3 \inst{2}}
\institute[Technische Universit\"at M\"unchen]
{ 
	%\inst{1} 
	Chair of IT Security (I20) \\Department of Informatics \\Technische Universit\"at M\"unchen 
	%\and
	%\inst{2} Cloud Service Lab\\Fraunhofer AISEC
}

@@

%% START PRESENTATION
\section{Fundamental of Causal Inference}
\subsection{Motivation}
\begin{frame}{What is Causality?}
\begin{block}{A definition from Wikipedia}
\textbf{Causality} (also referred to as causation) is the relationship
between an event (\textit{the cause}) and a second event (\textit{the effect}),
where the second event is understood as a consequence of the
first.
\end{block}\pause
\textbf{An example in real life} : Does smoking cause lung cancer?\\
\pause	
\begin{center}
\alert{Yes, it might be!}
\end{center}
\end{frame}
\begin{frame}{From Probabilistic View }
\textbf{Problem:} Does smoking cause lung cancer?\pause
\begin{center}
\begin{itemize}
\item Smoking does \alert{increase the probability} of getting lung cancer.
\end{itemize}
\end{center}
\end{frame}
\begin{frame}{Statistical Inference Overview}
\begin{figure}
\includegraphics[scale=0.7]{imgs/statInf}
\end{figure}
\begin{itemize}
\item Approximate an estimate of $X$ given evidence $e$, namely, $Pr(X\,|\,e)$. E.g., Regression or Classification problems.
\item Rejection of hypothesis, i.e., assert wether samples are from a certain distribution.  
\item Confidence interval, i.e., construct an interval based on dataset 
\end{itemize}
\end{frame}
\begin{frame}{Causal Inference Overview}
\begin{figure}
\includegraphics[scale=0.6]{imgs/causalInf}
\end{figure}
\begin{itemize}
\item What if $P$ has shifted itself to $P^*$?
\item<2-> \textbf{Key factors:} Causes, Changes, and Invariants . 
\item<2-> Inference of $P^*$ and reasoning of changes. 
\end{itemize}
\end{frame}
\begin{frame}{What makes Causal Inference interesting?}
\begin{itemize}
\item Human understands the world in terms of causes and effects.
\item Empirical science is about establishing causes.
\item Causal inference gives a mathematical language for causal
statements, and tools to solve causal problems formally.
\item Alternative exercising to decision making, reasoning, etc.
\end{itemize}
%\begin{alertblock}{Note}
%Causal Inference has fundamental difference with machine learning
%\end{alertblock}
\end{frame}
\subsection{Causal Graphical Model}
\begin{frame}{Association}
\begin{itemize}
\item Now we want to find out what \alert{causes} lung cancer
\end{itemize} \pause
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|}
\hline
& & \multicolumn{2}{|c|}{Lung cancer}\\\hline
smoking & yellow teeth & yes & no\\\hline
yes & yes & 100 & 400\\\hline
yes & no & 100 & 400\\\hline
no & yes & 1 & 450\\\hline
no & no & 9 & 8540\\\hline
\end{tabular}
\caption{Data observations from 10000 people}
\end{table}
\end{column}\hfill
\begin{column}{0.5\textwidth}
\vspace*{-0.5in}
\begin{figure}
\centering
\includegraphics[scale=0.6]{imgs/smokepl}
\end{figure}
\end{column}
\end{columns}
\end{frame}
\begin{frame}{Measurements of Association}
\textbf{To find out associations among variables}
\begin{itemize}
\item Mutual information (Information theory)
\item Pearson (linear) correlation
\item Spearman's rho (rank correlation)
\item Effect size between two variables
\item Many others
\end{itemize}
\end{frame}
\begin{frame}{Observations from Data}
\textbf{Obviously}
\begin{itemize}
\item \textit{yellow teeth} and \textit{lung cancer} are associated.
\end{itemize}\pause
\alert{\textbf{But...}}
\begin{itemize}
\item Bleaching the teeth does not help reduce the probability
of getting lung cancer.
\end{itemize}\pause
\begin{alertblock}{Caution!}
Correlation does not imply Causation
\end{alertblock}
\end{frame}
%% START NEW SECTION
\section{History}
